---
title: "Applied R in the Classroom"
author: "JD Long & Dusty Turner"
bibliography: bibliography.bib
output:
  pdf_document: default
  html_document:
    citation_package: biblatex
keep_tex: yes
link-citations: yes
csl: chicago-library-list.csl
biblio-style: authoryear
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




## Abstract

## Introduction - Dusty

The Father of Modern Econometrics, Adam Smith, once frustratingly stated:

> The discipline of colleges and universities is in general contrived, not for the benefit of the students, but for the... ease of the masters. [@smith]

<!-- > The discipline of colleges and universities is in general contrived, not for the benefit of the students, but for the interest, or more properly speaking, for the ease of the masters. -->

<!-- https://www.adamsmith.org/adam-smith-quotes -->
<!-- This article is in the lit review folder -->

Adam Smith is making the claim that methods of instruction used by professors are those which are easiest for the instructor, but not necessarily what is best for the student.  This is understandable, as research and other administive demands force instructors to rely on tried teaching techniques which are familiar to the teacher and not what will equip the student to be most successful in their career.

The authors of this article will make the case that the R open sourse statistical programming languge can bridge this gap between a Smith's proverbial teacher's ease and a student's benefit.  As R continues to be one of the more popular coding languages in industry with ever increasing technical support, the barriar for entry keeps falling.  There are many tools available in R which can aide the teaching process to get students loading and exploring data quickly requiring low overhead for proficiency to teach.  With ample open source support, a wide acceptance in industry and many additional features to explore and present data, teaching with R is both satisfies the ease for the instructor and has long term benefit for the students.  

This is something we are certain Adam Smith would be proud of.

<!-- Motivation to learn/teach r / Purpose of paper -->
<!-- Good thing is that there are all these addons -->


## The R Ecosystem - JD

### CRAN 

Describe Base R install, CRAN pacakges, and RStudio

### RStudio

R Studio & Rstudio.cloud - how to  share a project for the class

### Tidyverse

library(tidyverse) explain each package purpose

### Installation 

an aside on how to do this on your personal computer - link to R Cookbook

### Projects

Directory structure in the project. 


## Tutorial begins - Dusty will start


### Loading Data (Gapminder) 

```{r message=FALSE, warning=FALSE}
library(tidyverse)

gapminder <- read_csv("01_data/gapminder.csv")
```
Or directly from a URL:
will be a 404 until the repo is public

```{r eval=FALSE, message=FALSE, warning=FALSE}
gapminder <- read_csv("https://github.com/CerebralMastication/r_for_the_student/blob/master/01_data/gapminder.csv")
```


### Identifying data types

```{r, warning=FALSE}
summary(gapminder)
```

Or using the skimr package
```{r, warning=FALSE}
library(skimr)
skim(gapminder)

```


### `dplyr` Verbs 

(select, filter, mutate, group_by, summarize, arrange)


### Plotting/EDA
Esquisse? - https://cran.r-project.org/web/packages/esquisse/readme/README.html


### Regression 
parsnip/tidymodels
















### text from prior draft left below for salvaging

## An example of how to cite in R


With more than 275 practical recipes, this expanded edition [@teetor_long] helps you perform data analysis with R quickly and efficiently. The R language [@R-rmarkdown;@R-knitr] provides everything you need to do statistical work, but its structure can be difficult to master. These task-oriented recipes make you productive with R immediately. Solutions range from basic tasks to input and output, general statistics, graphics, and linear regression.

Each recipe addresses a specific problem and includes a discussion that explains the solution and provides insight into how it works. If you're a beginner, R Cookbook will help get you started. If you're an intermediate user, this book will jog your memory and expand your horizons @grolemund_wickham. You'll get the job done faster and learn more about R in the process.




#### R Studio Layout

One of the most common mistakes I run into in class on day one is for beginners to try to run R directly from the console.  They have R Studio downloaded, but do not understand the difference between R and R studio.  R is the program that executes all commands you provide.  It is the engine of the car, if you will.  R Studio is the shell that keeps the user organized.  R Studio is the shell around the car.  To continue, it is the steering wheel, accelerator, brakes, and seatbelts.  

To ensure you have opened R Studio, you screen should be broken into 3 to 4 panels.  

1) Top Right: The Environment
2) Bottom Right: The Files Viewer (what do you call this quadrants?)
3) Bottom Right: The Console
4) Top Left: The Script

Upon first opening R Studio, you may not have panel 4.  In order to open up a script, you will need to select `file` -> `New File` -> `R Script`.

Many new users also make the mistke of not realizing that they have not opened up a script.  They begin executing in the console again and do not take advantage of the ease that scripts provide in writing reproducable code.  

Possibly note on how to execute lines of code.  Both code chuncks and specific lines of code. 

### Tidyverse

## Installation

paraphrase what's in R Cookbook?

### Installing R

### Installing R Studio

### Installing the Tidyverse

## Using R

### Loading Data

### Plotting 

quick intro to ggplot

That graphical tool for ggplot?


### Plotting Multivariate Data

ggpairs?




<!-- When teaching, is the intent to teach coding in R, or is it to use R as an aide to teach statistics?  This will drive what you do and why you do it.   -->

<!-- ### What is the intent of teaching with R? -->

<!-- I have taught advanced introduction to probability and statistics and used R as a tool to teach statistics.  We introduced it by showing how it can be a good calculator -- and along the way taught some of the R Studio functionality (projects/setting working directory, how to execute lines of code, save variables, etc). -->

<!-- I have also taught sabermetrics where we taught coding in R.  So we needed to do data analysis (and we worked under the assumptions that our cadets knew basic statistics) so I taught the coding as a means of data analysis. -->

<!-- ### Do you "give stutents code"? -->

<!-- You have two options:  -->

<!-- 1) Give students code before class.   -->

<!-- The advantage here is that you can talk concepts and the syntax and inevitable mistyping, etc does not inhibit the larger purpose of the instruction.   -->

<!-- The disadvantage is that students don't get intimate with the coding process and learn the necessary skill of trial and error.   -->

<!-- 2) Don't give them code and expect them to keep up.   -->

<!-- The advantage is that cadets are learning the process of coding and "own" the code while they write it.   -->

<!-- The disadvantage is instead of teaching concepts, students are distracted by keeping up with the coding and missing out on the statistics lesson -->

<!-- My recommendation -- you give code if your intent is teaching statistics.  You don't give code if the intent is that cadets learn to code.   -->



# Bibliography

<!-- @grolemund_wickham -->
<!-- @long_teetor -->
<!-- @R-knitr -->
<!-- @R-rmarkdown -->